### 2022-05-31 paper.Rmd
minor editing

documentclass     : "apa6"
classoption       : "doc"
output            : 
  papaja::apa6_pdf: 
    number_sections: true
---

@@

Anticlustering is a collection of methods to partition a set of elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@feo1990]. Anticlustering thereby reverses the logic of its better known twin cluster analysis, which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis---including as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function, representing between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].

@@

The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively. Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1. When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as

@@

This formula again has the same form as $\mathit{SSE}$ (and $\mathit{SSE_{Var}}$), but is computed on different data. In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function. That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix. Hence, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.

During the remainder of this paper, whenever the k-plus criterion is optimized, the original data matrix is simply extended and the original k-means criterion $SSE$ is optimized. For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019]. In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending the state of the art on multicriterion optimization methods.

@@

To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.

>>>>

documentclass     : "apa6"
classoption       : "doc"
output            : 
  papaja::apa6_pdf
---

@@

Anticlustering is a collection of methods to partition elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@feo1990]. Anticlustering thereby reverses the logic of its better known twin cluster analysis, which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis---including as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function, representing between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].

@@

The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively. Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1. This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as

@@

This formula again has the same form as $\mathit{SSE}$ (and $\mathit{SSE_{Var}}$), but is computed on different data. In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function. That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix. Hence, using the unweighted sum approach, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.

@@

To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure. To implement all k-plus objectives, the simple unweighted sum outlined above approach was employed. Thus, when optimizing a k-plus criterion, the original data matrix was extended to incorporate the additional criteria. Subsequently, the original k-means criterion $SSE$ was maximized on the basis of the augmented data set.

### 2022-05-31 potpurri.Rmd
minor editing

>>>>

During the remainder of this paper, whenever the k-plus criterion is optimized, the original data matrix is simply extended and the original k-means criterion $SSE$ is optimized. For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019]. In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending the state of the art on multicriterion optimization methods.
